Pre-/Post-Trikafta in CFRD

### 2022-05-05 Colleen Wood/Pre-post Trikafta/pre_post_trikafta_analysis.Rmd
two years pre CFTR

```{r}
# Most recent before/after - limit to 2 years before start
# Want the value closest to start date before start, and most recent with respect to now from start
# (i.e. the longest time after start)
most_recent = df %>% group_by(CFF.ID,Site,Before_After) %>%
  filter(Time > -2*365) %>%
  mutate(time_new = max(Time) - Time) %>%
  arrange(time_new) %>% 
  summarise(Age = first(na.omit(Age)),BMI.Percentile = first(na.omit(BMI.Percentile)),
            Weight = first(na.omit(Weight)),Sex = first(na.omit(Sex)),
            T1D.Duration = first(na.omit(T1D.Duration)),
            G.tube.in.past.12.months = first(na.omit(G.tube.in.past.12.months)),
            Pancreatic.Status = first(na.omit(Pancreatic.Status)),
            Genotype1_sev = first(na.omit(Genotype1_sev)),
            Genotype2_sev = first(na.omit(Genotype2_sev)),
            BMIz = first(na.omit(BMIz)),Weightz = first(na.omit(Weightz)),
            FEV1 = first(na.omit(FEV1)),FVC = first(na.omit(FVC)),
            HbA1c = first(na.omit(HbA1c)),.groups = "keep") %>%
  arrange(Site,CFF.ID,Before_After)
# # Check
# t = df %>% select(CFF.ID,Site,Date,Time,Before_After,BMIz,ppFEV,ppFVC) %>%
#   group_by(CFF.ID,Site,Before_After) %>%
#   arrange(Site,CFF.ID,Before_After)
# Filter those without CFRD, and those without pre AND post
keep = most_recent %>% filter(!is.na(HbA1c)) %>%
  group_by(CFF.ID) %>% summarise(n = n()) %>%
  filter(n == 2) %>% pull(CFF.ID)
cfrd = unique(df$CFF.ID[!is.na(df$CFRD.Status) | !is.na(df$CFRD.Diagnosis.Date)])
keep = unique(keep,cfrd)
most_recent = most_recent %>% filter(CFF.ID %in% keep)
```

>>>>

```{r}
# Most recent before/after - limit to 2 years before start
# Want the value closest to start date before start, and most recent with respect to now from start
# (i.e. the longest time after start)
most_recent = df %>% group_by(CFF.ID,Site,Before_After) %>%
  filter(Time >= (-2*365)) %>%
  mutate(time_new = max(Time) - Time) %>%
  arrange(time_new) %>% 
  summarise(Start = first(na.omit(Start)),
            Age = paste0(Date[which(Age == first(na.omit(Age)))],": ",round(first(na.omit(Age)),1)),
            BMI.Percentile = paste0(Date[which(BMI.Percentile == first(na.omit(BMI.Percentile)))],
                                    ": ",round(first(na.omit(BMI.Percentile)),1)),
            Weight = paste0(Date[which(Weight == round(first(na.omit(Weight)),1))],
                            ": ",first(na.omit(Weight))),
            T1D.Duration = paste0(Date[which(T1D.Duration == first(na.omit(T1D.Duration)))],
                                  ": ",round(first(na.omit(T1D.Duration)),1)),
            G.tube.in.past.12.months = first(na.omit(G.tube.in.past.12.months)),
            Pancreatic.Status = first(na.omit(Pancreatic.Status)),
            Genotype1_sev = first(na.omit(Genotype1_sev)),
            Genotype2_sev = first(na.omit(Genotype2_sev)),
            BMIz = paste0(Date[which(BMIz == first(na.omit(BMIz)))],
                          ": ",round(first(na.omit(BMIz)),1)),
            FEV1 = paste0(Date[which(FEV1 == first(na.omit(FEV1)))[1]],
                          ": ",round(first(na.omit(FEV1)),1)),
            FVC = paste0(Date[which(FVC == first(na.omit(FVC)))],
                         ": ",round(first(na.omit(FVC)),1)),
            HbA1c = paste0(Date[which(HbA1c == first(na.omit(HbA1c)))],
                           ": ",round(first(na.omit(HbA1c)),1)),
            .groups = "keep") %>%
  arrange(Site,CFF.ID,Before_After)
most_recent$Start = as.character(most_recent$Start)
most_recent[most_recent == ": NA"] = NA
most_recent[most_recent == "NA: NA"] = NA

# # Check
# t = df %>% select(CFF.ID,Site,Date,Time,Before_After,BMIz,ppFEV,ppFVC) %>%
#   group_by(CFF.ID,Site,Before_After) %>%
#   arrange(Site,CFF.ID,Before_After)
# Filter those without CFRD, and those without pre AND post
keep = most_recent %>% filter(!is.na(HbA1c)) %>%
  group_by(CFF.ID) %>% summarise(n = n()) %>%
  filter(n == 2) %>% pull(CFF.ID)
cfrd = unique(df$CFF.ID[!is.na(df$CFRD.Status) | 
                          df$CFRD.Status != "Impaired Glucose Tolerance (FBG < 126, 2-h PG 140-199)" | 
                          !is.na(df$CFRD.Diagnosis.Date)])
keep = unique(keep,cfrd)
most_recent = most_recent %>% filter(CFF.ID %in% keep)
# Write 
write.csv(most_recent,file = "./Colleen Wood/Prepost Triakfta/Data_Cleaned/pre_post.csv",
          na = "",row.names = F)
```
